The Birth of a Galaxy: Primordial Metal Enrichment and Stellar Populations

### 2011-11-30 ApJ/ms_2c.tex
Updating emulateapj version with the accepted version.

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$ \citep{ABN02, Bromm02_P3, OShea07a,
  Yoshida08}.  They form in molecular clouds that coexist with the
dark matter halo center with limited fragmentation occurring during
their collapse; however \citet{2009Sci...325..601T} and
\citet{Stacy10_Binary} have shown that Pop III binaries may form in a
fraction of such halos.  There have been some recent hints that the
accretion disk around the first protostar may fragment and produce
clustered Pop III star formation \citep{Clark11_Frag,
  Greif11_P3Cluster}.  The final stellar masses in the aforementioned
calculations are still highly uncertain with a limited number of
samples.  Nevertheless, they all point toward an IMF with a larger
characteristic mass than present-day star formation.

@@

@@

We have focused on the birth of two galaxies prior to reionization
with a cosmological AMR radiation hydrodynamics simulation with
unprecedented spatial (1 pc comoving) and mass resolution (1800 \Ms).
With this resolution, we are able to follow the Pop III star formation
and radiative and chemical feedback effects in hundreds of minihalos
that contribute to the formation of a protogalaxy.  Supernovae from
Pop III stars provide the necessary heavy elements for the transition
to a Population II stellar population, which we have directly
simulated.  These two galaxies have a 10--15\% probability in
surviving as present-day ``fossil'' galaxies \citep{Gnedin06},
otherwise they will be incorporated into galactic stellar halos.  A
$z=8.55$ galaxy was recently spectroscopically confirmed that is
contained in an ionized bubble with radius 0.1--0.5 Mpc with an
uncharacterized population of galaxies contributing to the local
ionizing radiation field \citep{Lehnert10_z8.6}.  The galaxies
simulated here might represent this population and should be detected
with the upcoming \textit{James Webb Space Telescope}.

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given $M_{\rm char} = 100~\Ms$.  DLA systems have a metallicity floor
on the same order \citep{Wolfe05_Review, Penprase10}, and metal
enrichment from Pop III SNe provides a possible explanation
\citep{Kobayashi11}.  Ours is the first simulation with sufficient
physics and resolution to support this claim.

@@

In the case where this metallicity floor is less than the critical
metallicity, then the next SN will further enrich the ISM, solidifying
the transition to Pop II SF.  We conclude that it only takes one, at
most two, SNe from Pop III stars in the halo progenitors to complete
the transition to Population II \citep{Frebel10}.  However, this does
not prohibit more than two instances of Pop III star formation in halo
from happening because the stars can end in a direct collapse to a BH
with little or no metal ejecta, keeping the surrounding environment
pristine.  The question of whether the critical metallicity is
\tento{-6} or \tento{-3.5} $Z_\odot$ is most applicable to nearby
halos where the heavy elements mix slowly into dense cores as the
blastwave overtakes it.  Less than two percent of stars have [Z/H] $<
-3$ in both systems, consistent with observations of metal-poor stars
in the halo and dSphs \citep{Beers05, Battaglia11}.

@@

JHW thanks Renyue Cen, Amina Helmi, Marco Spaans, and Eline Tolstoy
for enlightening discussions.  JHW is partially supported by NASA
through Hubble Fellowship grant \#120-6370 awarded by the Space
Telescope Science Institute, which is operated by the Association of
Universities for Research in Astronomy, Inc., for NASA, under contract
NAS 5-26555.  MJT and MLN acknowledge partial support from NASA ATFP
grant NNX08AH26G.  Computational resources were provided by NASA/NCCS
award SMD-09-1439.  TA acknowledges financial support from the {\em
  Baden-W\"{u}rttemberg-Stiftung} under grant P-LS-SPII/18, the
Heidelberg Institut f\"ur Theoretische Studien. This work was
partially supported by NASA ATFP grant NNX08AH26G, NSF AST-0807312,
and NSF AST-1109243. This research has made use of NASA’s Astrophysics
Data System Bibliographic Services.  The majority of the analysis and
plots were done with \texttt{yt} \citep{yt_full_paper}.

%\clearpage
\begin{thebibliography}{101}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

@@

\bibitem[{{Bovill} \& {Ricotti}(2010{\natexlab{a}})}]{Bovill11a}
{Bovill}, M.~S., \& {Ricotti}, M. 2010{\natexlab{a}}, ArXiv e-prints
  (1010.2231)

\bibitem[{{Bovill} \& {Ricotti}(2010{\natexlab{b}})}]{Bovill11b}
---. 2010{\natexlab{b}}, ArXiv e-prints (1010.2233)

@@

@@

\bibitem[{{Gnedin} \& {Kravtsov}(2006)}]{Gnedin06}
{Gnedin}, N.~Y., \& {Kravtsov}, A.~V. 2006, \apj, 645, 1054

@@

@@

@@

@@

@@

@@

\bibitem[{{Tseliakhovich} {et~al.}(2010){Tseliakhovich}, {Barkana}, \&
  {Hirata}}]{Tselia10_Minihalo}
{Tseliakhovich}, D., {Barkana}, R., \& {Hirata}, C. 2010, ArXiv e-prints

@@

@@

\bibitem[{{Wolcott-Green} {et~al.}(2011){Wolcott-Green}, {Haiman}, \&
  {Bryan}}]{Wolcott11}
{Wolcott-Green}, J., {Haiman}, Z., \& {Bryan}, G.~L. 2011, ArXiv e-prints

@@

>>>>

\subsubsection{Population III star formation}

Simulations following the full accretion history of Pop III stars,
from their formation inside cosmological minihalos through
transitioning onto the main sequence, are not currently possible.
However, under the assumption of spherical accretion, reasonable
estimates of the mass ranges for Pop III stars can be made by
extrapolating forward in time the instantaneous accretion rate at the
time of formation of the first protostellar core and comparing the
timescales for accretion against the Kelvin-Helmholtz time.  Several
studies, conducting this extrapolation, have shown that the
characteristic mass of Pop III stars $M_{\rm char} \sim 100~\Ms$
\citep{ABN02, Bromm02_P3, OShea07a, Yoshida08}.  They form in
molecular clouds that coexist with the dark matter halo center with
limited fragmentation occurring during their collapse; however
\citet{2009Sci...325..601T} and \citet{Stacy10_Binary} have shown that
Pop III binaries may form in a fraction of such halos.  There have
been some recent hints that the accretion disk around the first
protostar may fragment and produce clustered Pop III star formation
\citep{Clark11_Frag, Greif11_P3Cluster}.  The final stellar masses in
the aforementioned calculations are still highly uncertain with a
limited number of samples.  Nevertheless, they all point toward an IMF
with a larger characteristic mass than present-day star formation.

@@

\begin{figure*}
\epsscale{1.15}
\plotone{f1}
\epsscale{1}
  \caption{\label{fig:evo-mosaic} Evolution of the entire simulation
    volume ($L_{\rm box} = 1$ Mpc) at redshifts 15, 12, 10, 8, and 7
    (left to right).  Pictured here are the density-weighted
    projections of density (top), temperature (middle), and
    metallicity (bottom).  Note how the stellar radiative feedback
    from low-mass galaxies reionize the majority of the volume.  The
    metallicity projections are a composite image of metals
    originating from Pop II (red) and III (blue) stars with magenta
    indicating a mixture of the two.}
\end{figure*}

@@

We have focused on the birth of two galaxies prior to reionization
with a cosmological AMR radiation hydrodynamics simulation with very
high spatial (1 pc comoving) and mass resolution (1800 \Ms).  With
this resolution, we are able to follow the Pop III star formation and
radiative and chemical feedback effects in hundreds of minihalos that
contribute to the formation of a protogalaxy.  Supernovae from Pop III
stars provide the necessary heavy elements for the transition to a
Population II stellar population, which we have directly simulated.
The galaxies simulated here represent a population of uncharacterized
high-redshift dwarf galaxies, which should be detected with the
upcoming \textit{James Webb Space Telescope}, that should contribute a
significant fraction of the ionizing photons during reionization.
Recently, a $z=8.6$ spectroscopically confirmed galaxy was found to
reside in an \ion{H}{2} region that extends beyond 1 Mpc.  However,
using the SFRs and \lya~continuum as estimators, this galaxy produces
enough ionizing radiation to create an \ion{H}{2} region between $\sim
0.1(f_{\rm esc}/0.1)^{1/3}$ Mpc and $0.5(f_{\rm esc}/0.1)^{1/3}$ Mpc,
where $f_{\rm esc}$ is the fraction of ionizing radiation escaping
into the IGM \citep{Lehnert10_z8.6}.  This discrepancy could be caused
by undetected satellite dwarf galaxies, such ones presented here, that
contribute to local radiation field, or an ionizing radiation escape
fraction near unity \citep{Wise09, Razoumov10, Paardekooper11,
  Yajima11}.

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given $M_{\rm char} = 100~\Ms$.  DLA systems have a metallicity floor
on the same order \citep{Wolfe05_Review, Penprase10}, and metal
enrichment from Pop III SNe provides a possible explanation
\citep{Kobayashi11}.  Our simulation strengthens this claim of a
metallicity floor from Pop III SNe \citep[see also][]{Tornatore07,
  Karlsson08, Maio11_Enrich}, where our work improves on previous
studies by resolving star-forming molecular clouds and employing more
realistic feedback physics, in particular, radiation transport.

@@

Building upon this idea of a lower metallicity floor, consider a
pre-galactic halo that has hosted a Pop III star, producing a SN.  The
majority of the gas is ejected into the IGM and re-accretes after
$\sim 10-50$ Myr \citep{Johnson07, Wise08_Gal}.  What happens if the
metallicity is not greater than the critical metallicity?  Massive Pop
III stars would form again in this halo; however, enriching the
surrounding IGM with Pop III SNe and ejecting material becomes more
difficult in halos with $M \gsim 10^7 \Ms$ \citep{Whalen08_SN}.  Thus
it is more likely for these halos to retain more of its gas during
star formation episodes and self-enrich its host halo and less of the
IGM.  This would accelerate the transition from Pop III to II in more
massive halos.

We conclude that it only takes one, at most two, SNe from Pop III
stars in the halo progenitors to complete the transition to Population
II \citep{Frebel10}.  However, this does not prohibit more than two
instances of Pop III star formation in halo from happening because the
stars can produce BHs with little or no metal ejecta, keeping the
surrounding environment pristine.  Even though we find that Pop III
SNe create a floor of $\sim 10^{-3} \zsun$, gas with lower
metallicities can still exists in nearby halos, where the metal-rich
SN ejecta will mix slowly into dense cores as the blastwave overtakes
them \citep{Cen08}.  These environments may be the host of true
extremely metal-poor star formation, such as the recently discovered
halo star with all detected metals below $10^{-4.2} \zsun$
\citep{Caffau11}.

@@

JHW thanks Renyue Cen, Amina Helmi, Marco Spaans, and Eline Tolstoy
for enlightening discussions.  We thank an anonymous referee for
helpful suggestions.  JHW is partially supported by NASA through
Hubble Fellowship grant \#120-6370 awarded by the Space Telescope
Science Institute, which is operated by the Association of
Universities for Research in Astronomy, Inc., for NASA, under contract
NAS 5-26555.  MJT and MLN acknowledge partial support from NASA ATFP
grant NNX08AH26G.  MJT~acknowledges support by the NSF CI TraCS
fellowship award OCI-1048505.  Computational resources were provided
by NASA/NCCS award SMD-09-1439.  TA acknowledges financial support
from the {\em Baden-W\"{u}rttemberg-Stiftung} under grant
P-LS-SPII/18, the Heidelberg Institut f\"ur Theoretische Studien. This
work was partially supported by NASA ATFP grant NNX08AH26G, NSF
AST-0807312, and NSF AST-1109243. This research has made use of NASA’s
Astrophysics Data System Bibliographic Services.  The majority of the
analysis and plots were done with \texttt{yt} \citep{yt_full_paper}.

%\clearpage
%\bibliography{ms}
\begin{thebibliography}{108}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

@@

\bibitem[{{Bovill} \& {Ricotti}(2011{\natexlab{a}})}]{Bovill11a}
{Bovill}, M.~S., \& {Ricotti}, M. 2011{\natexlab{a}}, \apj, 741, 17

\bibitem[{{Bovill} \& {Ricotti}(2011{\natexlab{b}})}]{Bovill11b}
---. 2011{\natexlab{b}}, \apj, 741, 18

@@

\bibitem[{{Caffau} {et~al.}(2011){Caffau}, {Bonifacio}, {Fran{\c c}ois},
  {Sbordone}, {Monaco}, {Spite}, {Spite}, {Ludwig}, {Cayrel}, {Zaggia},
  {Hammer}, {Randich}, {Molaro}, \& {Hill}}]{Caffau11}
{Caffau}, E., {Bonifacio}, P., {Fran{\c c}ois}, P., {Sbordone}, L., {Monaco},
  L., {Spite}, M., {Spite}, F., {Ludwig}, H.-G., {Cayrel}, R., {Zaggia}, S.,
  {Hammer}, F., {Randich}, S., {Molaro}, P., \& {Hill}, V. 2011, \nat, 477, 67

@@

@@

\bibitem[{{Johnson} {et~al.}(2007){Johnson}, {Greif}, \& {Bromm}}]{Johnson07}
{Johnson}, J.~L., {Greif}, T.~H., \& {Bromm}, V. 2007, \apj, 665, 85

@@

\bibitem[{{Martin} {et~al.}(2008){Martin}, {de Jong}, \& {Rix}}]{Martin08_uFd}
{Martin}, N.~F., {de Jong}, J.~T.~A., \& {Rix}, H.-W. 2008, \apj, 684, 1075

@@

\bibitem[{{Paardekooper} {et~al.}(2011){Paardekooper}, {Pelupessy}, {Altay}, \&
  {Kruip}}]{Paardekooper11}
{Paardekooper}, J.-P., {Pelupessy}, F.~I., {Altay}, G., \& {Kruip}, C.~J.~H.
  2011, \aap, 530, A87

@@

\bibitem[{{Razoumov} \& {Sommer-Larsen}(2010)}]{Razoumov10}
{Razoumov}, A.~O., \& {Sommer-Larsen}, J. 2010, \apj, 710, 1239

@@

\bibitem[{{Springel} \& {Hernquist}(2003)}]{Springel03b}
{Springel}, V., \& {Hernquist}, L. 2003, \mnras, 339, 289

@@

\bibitem[{{Tseliakhovich} {et~al.}(2011){Tseliakhovich}, {Barkana}, \&
  {Hirata}}]{Tselia10_Minihalo}
{Tseliakhovich}, D., {Barkana}, R., \& {Hirata}, C.~M. 2011, \mnras, 1501

@@

\bibitem[{{Whalen} {et~al.}(2008){Whalen}, {van Veelen}, {O'Shea}, \&
  {Norman}}]{Whalen08_SN}
{Whalen}, D., {van Veelen}, B., {O'Shea}, B.~W., \& {Norman}, M.~L. 2008, \apj,
  682, 49

@@

\bibitem[{{Wolcott-Green} {et~al.}(2011){Wolcott-Green}, {Haiman}, \&
  {Bryan}}]{Wolcott11}
{Wolcott-Green}, J., {Haiman}, Z., \& {Bryan}, G.~L. 2011, \mnras, 1673

@@

\bibitem[{{Yajima} {et~al.}(2011){Yajima}, {Choi}, \& {Nagamine}}]{Yajima11}
{Yajima}, H., {Choi}, J.-H., \& {Nagamine}, K. 2011, \mnras, 412, 411
